Preliminary Thesis Defence

### 2013-12-28 bitremit.tex
break away chapters

\documentclass[a4paper,twopage]{report}\usepackage{graphicx, color}
%% maxwidth is the original width if it is less than linewidth
%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

\definecolor{fgcolor}{rgb}{0.2, 0.2, 0.2}
\newcommand{\hlnumber}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hlfunctioncall}[1]{\textcolor[rgb]{0.501960784313725,0,0.329411764705882}{\textbf{#1}}}%
\newcommand{\hlstring}[1]{\textcolor[rgb]{0.6,0.6,1}{#1}}%
\newcommand{\hlkeyword}[1]{\textcolor[rgb]{0,0,0}{\textbf{#1}}}%
\newcommand{\hlargument}[1]{\textcolor[rgb]{0.690196078431373,0.250980392156863,0.0196078431372549}{#1}}%
\newcommand{\hlcomment}[1]{\textcolor[rgb]{0.180392156862745,0.6,0.341176470588235}{#1}}%
\newcommand{\hlroxygencomment}[1]{\textcolor[rgb]{0.43921568627451,0.47843137254902,0.701960784313725}{#1}}%
\newcommand{\hlformalargs}[1]{\textcolor[rgb]{0.690196078431373,0.250980392156863,0.0196078431372549}{#1}}%
\newcommand{\hleqformalargs}[1]{\textcolor[rgb]{0.690196078431373,0.250980392156863,0.0196078431372549}{#1}}%
\newcommand{\hlassignement}[1]{\textcolor[rgb]{0,0,0}{\textbf{#1}}}%
\newcommand{\hlpackage}[1]{\textcolor[rgb]{0.588235294117647,0.709803921568627,0.145098039215686}{#1}}%
\newcommand{\hlslot}[1]{\textit{#1}}%
\newcommand{\hlsymbol}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hlprompt}[1]{\textcolor[rgb]{0.2,0.2,0.2}{#1}}%

\usepackage{framed}
\makeatletter
\newenvironment{kframe}{%
 \def\at@end@of@kframe{}%
 \ifinner\ifhmode%
  \def\at@end@of@kframe{\end{minipage}}%
  \begin{minipage}{\columnwidth}%
 \fi\fi%
 \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
 \colorbox{shadecolor}{##1}\hskip-\fboxsep
     % There is no \\@totalrightmargin, so:
     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
 \MakeFramed {\advance\hsize-\width
   \@totalleftmargin\z@ \linewidth\hsize
   \@setminipage}}%
 {\par\unskip\endMakeFramed%
 \at@end@of@kframe}
\makeatother

\definecolor{shadecolor}{rgb}{.97, .97, .97}
\definecolor{messagecolor}{rgb}{0, 0, 0}
\definecolor{warningcolor}{rgb}{1, 0, 1}
\definecolor{errorcolor}{rgb}{1, 0, 0}
\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX

\usepackage{alltt}
\usepackage{mathtools}
\usepackage{rotating}
\usepackage{graphicx}
\usepackage[authordate,backend=biber]{biblatex-chicago}
\usepackage[colorlinks=true,linkcolor=black,citecolor=black,urlcolor=black]{hyperref}

\title{Preliminary Thesis Defence\\~\\
\begin{tabular}{rl}
Supervisor:&Jean-Louis Arcand\footnote{Professor of Economics, The Graduate Institute, Geneva; jean-louis.arcand@graduateinstitute.ch}\\
Second Reader:&Ugo Panizza\footnote{Professor of Economics, The Graduate Institute, Geneva; ugo.panizza@graduateinstitute.ch}
\end{tabular}
}

\author{Bastiaan Quast\thanks{PhD Student, The Graduate Institute, Geneva; bastiaan.quast@graduateinstitute.ch / bquast@gmail.com}\\~\\
The Graduate Institute, Geneva}

\let\oldmarginpar\marginpar
\renewcommand\marginpar[1]{\-\oldmarginpar[\raggedleft\footnotesize #1]
{\raggedright\footnotesize #1}}

\addglobalbib{bibliography.bib}

\newbibmacro{string+doiurlisbn}[1]{%
  \iffieldundef{doi}{%
    \iffieldundef{url}{%
      \iffieldundef{isbn}{%
        \iffieldundef{issn}{%
          #1%
        }{%
          \href{http://books.google.com/books?vid=ISSN\thefield{issn}}{#1}%
        }%
      }{%
        \href{http://books.google.com/books?vid=ISBN\thefield{isbn}}{#1}%
      }%
    }{%
      \href{\thefield{url}}{#1}%
    }%
  }{%
    \href{http://dx.doi.org/\thefield{doi}}{#1}%
  }%
}

\DeclareFieldFormat{title}{\usebibmacro{string+doiurlisbn}{\mkbibemph{#1}}}
\DeclareFieldFormat[article,incollection]{title}%
    {\usebibmacro{string+doiurlisbn}{\mkbibquote{#1}}}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}

\begin{document}

\maketitle
\tableofcontents
\listoftables
\listoffigures

\begin{refsection}
\chapter{Introduction}
In this Preliminary Thesis Defence I present the research projects that I intend to use for my PhD dissertation. These projects are titled:

\begin{itemize}
\item Additional South African Evidence on Pensions and Child Growth (\ref{south-africa})
\item Replication: Surviving Andersonville (\ref{andersonville})
\item Risk and Uncertainty in Unique Equilibria in Currency Attacks (\ref{currency})
\item Bitcoin and Remittances: The Potential of `Stupid Phones' (\ref{bitcoin})
\end{itemize}

These project are in various stages of completion. Below I will briefly describe the research idea, the current state, and the potential for each project.

Firstly, the paper Additional South African Evidence of Pensions and Child Growth builds \textcite{duflo2000, duflo2003}. These papers look at the effect of the gender of grandparents on the growth metrics of their grandchildren. A problem in this comparison lies in the fact that pension eligibility of men was at 65 years, whereas it was 60 for women. This paper makes use of a Household Survey from 2008, 2012 and 2013. The survey is especially of interest, because the pension eligibility age for men was lowered from 65 to 60 in 2009. As it stands now I have analysed the data from 2008 and 2012. The 2013 data was released several weeks ago, I will incorporate these data in future versions of my analysis.

Secondly, together with several other PhD students we have replicated the paper ``\citetitle{costa2007surviving}'' by \textcite{costa2007surviving}. We were able to replicate all results from the paper, save for a few decimal numbers, which we have taken to be the result of rounding errors. However, replicating these numbers required us to take a number of unorthodox steps in our analysis. When in stead, we take more orthodox steps, we find that we cannot quantitatively and qualitatively replicate the authors' results. As an example, we find that the dataset distinguishes between two different POW camps both located in the town of Andersonville. The authors however, aggregate these two. This leads to overstated magnitudes of the social ties of POWs. This is problematic because the main finding of the paper is that the size of these social networks is relevant to an individual's odds of survival. At this point our replication is quite complete. Before rewriting this paper, I will redo the analysis in R \parencite{R}, currently the original and our replication are in Stata.

Thirdly, \textcite{morris1998unique} claims that introducing a measure of risk in a model of currency attacks, leads to a unique equilibrium. It thus solves the previous multiple-equilibrium zone called `ripe for attack' for a unique equilibrium. I believe that this is a result of the fact that the risk factor introduces replaces uncertainty, not certainty. We thus gain information. I want to discuss MS1998 in light of the distinction between Knightian uncertainty and quantified risk.

Fourthly, Bitcoin is the first of a new breed of currencies, commonly referred to as a cryptocurrencies. Bitcoin transactions are conducted using an electronic process called signing. I would like to study the role of Bitcoin in remittances. For this I would like to participate in building Bitcoin applications for the simple type of phones common in developing economies (sometimes referred to stupid phones). We could study the spread of bitcoin by looking at the downloads of the apps from regions, as well looking at how releasing the app in different languages evolves.

After this the papers follow, in the order listed above. A separate citations section is included for each chapter. For ease of reference, the titles of the citations in the bibliography link to the online articles using DOIs and URLs. In the end notes (\ref{end-notes}) I present a brief discussion of my dissertation, as it stands now, a whole. I also mention some very preliminary ideas.

\begin{refsection}
\chapter{Pensions and Child Growth in South Africa}
\label{south-africa}
\section*{Abstract}

In this paper I look at the effect of the gender of pension recipients on the growth of children in the same households.

I do this by comparing z-scores of anthropometrics of South African children living in the same household as state pension recipients. This paper exploits the fact that the data set consists of two surveys which were done before and after the lowering of the pension eligibility age for men to the same age as women.

The main preliminary finding is that the household effect was only significant in 2012.

\section{Introduction}

This paper looks at the effect of the gender of pension recipients on the growth of children in their household in South Africa. The approach
is very similar to \textcite{duflo2000, duflo2003}. The deviation from international standards \parencite{who2006child} for weight-for-height and length-for-age are computed as z-scores. These are then compared for different pension recipient status and gender.

The anthropometrics are useful for computing z-scores. These z-scores are considered a good representation of short-term or long-term malnutrition respectively, especially for children between 6 and 60 months old.

The South African pension is a interesting variable to measure income because of its criteria. Besides a maximum level of income, the only criterium is the age of a person. Because of this the status as a recipient is quite exogenous and there are few selection bias issues. The problematic difference between the eligibility age of men and women was eliminated between the two surveys which creates an interesting natural experiment.

Duflo finds evidence in South Africa's 1993 \href{http://microdata.worldbank.org/index.php/catalog/297}{Integrated Household Survey} that girls' short term nutrition (weight-for-height) was positively influence by living with the maternal grandmother if the grandmother was eligible for a state pension. This is directly after the significant increase in the pension sum for blacks. The pension-eligibility age at this time was 60 for women and 65 for men.

This study deviates from the Duflo study in several ways. In 2008 and 2012, the first and second wave of the \textcite{saldru2008nids, saldru2012nids} have gathered similar data. In the period 2008-2010, the government lowered the eligibility age for men from 65 to 60 (\href{http://www.southafrica.info/services/government/pension-160708.htm}{Announcement}).

This happened in a few steps. As of July 16th 2008 men aged 63 and 64 were eligible for a pension. Men aged 61 and 62 became eligible in April 2009. Finally in January 2010, pension eligibility age was at 60 for all citizens.

Another deviation is the usage of WHO z-scores \parencite{who2006child} in stead of CDC z-scores \parencite{nchs2000cdc}, since these have superseded the CDC charts, however this should not be of any consequence, as long as we use the same standards consistently.

The main preliminary result is that the household effect (i.e.~having one or more recipients in the household) is very insignificant in the 2008 estimate. But in the 2012 estimate it is very significant.

\section{Data}

In this paper I use data from two sources. The first is the South African National Income Dynamics Survey \parencite{saldru2008nids, saldru2012nids} and the second is the World Health Organization's Child Growth Standards \parencite{who2006child}.

The main source of data is the NIDS. This survey collects data on a representative set of appproximately 10,000 South African households. The primary information variables that I use are, the child anthropometrics, the age and gender of household members, and the status as state pension recipients.

For adults, several variables measure the different amounts and sources of income. Among those, a variable for the adult receiving a state pension, and if so, how much. This is a numeric variable, the values of which lie very close together. For simplicity I have temporarily used this variable as a dummy.

The children's anthropometrics taken are, length/height, weight, and waist. Using these anthropometics and WHO growth standards, z-scores have been calculated. Unfortunately, the latest version of wave 2 (2012) accidentally omitted the z-scores, so that these cannot be evaluated until an updated version is published. However, I have computed the length-for-age z-scores manually.

In 2006 the WHO published its standards for child growth. These standards are based on the scores of children from different ethnic populations in households which maintained a healthy lifestyle. The standards provide the means and standards deviations used. These are on a monthly basis for height-for-age, and on a semi-centimeter level for weight-for-height scores.

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
p1 <- \hlfunctioncall{ggplot}(\hlfunctioncall{subset}(w1_child, w1_woman_60 == 1))
p1 <- p1 + \hlfunctioncall{aes}(w1_c_age_m, w1_zhfa, colour = \hlfunctioncall{factor}(w1_c_woman))
p1 <- p1 + \hlfunctioncall{stat_smooth}(method = \hlstring{"loess"}) + \hlfunctioncall{geom_point}()
p1
\end{alltt}

{\ttfamily\noindent\color{warningcolor}{\#\# Warning: Removed 324 rows containing missing values (stat\_smooth).}}

{\ttfamily\noindent\color{warningcolor}{\#\# Warning: Removed 263 rows containing missing values (stat\_smooth).}}

{\ttfamily\noindent\color{warningcolor}{\#\# Warning: Removed 51 rows containing missing values (stat\_smooth).}}

{\ttfamily\noindent\color{warningcolor}{\#\# Warning: Removed 638 rows containing missing values (geom\_point).}}\end{kframe}
\includegraphics[width=\maxwidth]{figure/r-p1}

\end{knitrout}

\section{Results}

Basic HAZ (2008)

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlfunctioncall{summary}(w1_haz0)
\end{alltt}
\begin{verbatim}
## 
## Call:
## lm(formula = w1_zhfa ~ w1_spen_w + w1_spen_m + w1_h_tinc + w1_best_edu + 
##     w1_best_age_yrs, data = w1_child, subset = w1_c_age_m >= 
##     6 & w1_c_age_m <= 60)
## 
## Residuals:
##    Min     1Q Median     3Q    Max 
## -4.779 -1.036 -0.169  0.834  7.060 
## 
## Coefficients:
##                  Estimate Std. Error t value Pr(>|t|)    
## (Intercept)     -1.43e+00   1.44e-01   -9.87   <2e-16 ***
## w1_spen_w        2.32e-02   1.15e-01    0.20   0.8402    
## w1_spen_m        1.36e-01   1.72e-01    0.79   0.4281    
## w1_h_tinc        3.62e-05   9.99e-06    3.62   0.0003 ***
## w1_best_edu      1.63e-03   7.84e-03    0.21   0.8355    
## w1_best_age_yrs  7.61e-03   3.64e-03    2.09   0.0369 *  
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## 
## Residual standard error: 1.77 on 1436 degrees of freedom
##   (1936 observations deleted due to missingness)
## Multiple R-squared:  0.0134,	Adjusted R-squared:  0.00994 
## F-statistic: 3.89 on 5 and 1436 DF,  p-value: 0.00164
\end{verbatim}
\end{kframe}
\end{knitrout}

HAZ 2008, with 2012 eligiblity as an IV
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlfunctioncall{summary}(tsls_w1w2_haz0)
\end{alltt}
\begin{verbatim}
## 
##  2SLS Estimates
## 
## Model Formula: w1_zhfa ~ w1_spen_w + w1_spen_m + w1_h_tinc + w1_best_edu + w1_best_age_yrs
## 
## Instruments: ~w2_spen_w + w2_spen_m + w1_h_tinc + w1_best_edu + w1_best_age_yrs
## 
## Residuals:
##    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
##  -4.990  -0.835  -0.050   0.000   0.788   7.070 
## 
##                   Estimate Std. Error t value Pr(>|t|)    
## (Intercept)     -1.172e+00  7.388e-02 -15.861  < 2e-16 ***
## w1_spen_w        1.748e-02  9.477e-02   0.184   0.8536    
## w1_spen_m       -1.374e-01  1.444e-01  -0.952   0.3414    
## w1_h_tinc        2.995e-05  5.606e-06   5.342 9.69e-08 ***
## w1_best_edu      2.410e-03  3.532e-03   0.682   0.4951    
## w1_best_age_yrs  4.127e-03  1.863e-03   2.215   0.0268 *  
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## 
## Residual standard error: 1.4889 on 4227 degrees of freedom
\end{verbatim}
\end{kframe}
\end{knitrout}

HAZ 2008, with 2012 eligiblity as an IV (men over 65 only)
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlfunctioncall{summary}(tsls_w1w2_haz1)
\end{alltt}
\begin{verbatim}
## 
##  2SLS Estimates
## 
## Model Formula: w1_zhfa ~ w1_spen_w + w1_spen_m + w1_h_tinc + w1_best_edu + w1_best_age_yrs
## 
## Instruments: ~w2_spen_w + w2_spen_m_65 + w1_h_tinc + w1_best_edu + w1_best_age_yrs
## 
## Residuals:
##    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
##  -4.990  -0.836  -0.050   0.000   0.786   7.040 
## 
##                   Estimate Std. Error t value Pr(>|t|)    
## (Intercept)     -1.173e+00  7.387e-02 -15.880  < 2e-16 ***
## w1_spen_w        8.576e-03  9.541e-02   0.090   0.9284    
## w1_spen_m       -9.189e-02  1.348e-01  -0.681   0.4956    
## w1_h_tinc        2.992e-05  5.605e-06   5.339 9.85e-08 ***
## w1_best_edu      2.443e-03  3.531e-03   0.692   0.4890    
## w1_best_age_yrs  4.092e-03  1.860e-03   2.200   0.0278 *  
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## 
## Residual standard error: 1.4885 on 4227 degrees of freedom
\end{verbatim}
\end{kframe}
\end{knitrout}

\printbibliography

\end{refsection}

\begin{refsection}
\chapter{Replication: Surviving Andersonville}
\label{andersonville}
\textbf{Aksel Erbahar, Cecilia Heuser, Victor K\"ummritz, Bastiaan Quast}

\section*{Abstract}
We replicate \textcite{costa2007surviving}. We are able to replicate all the result from the paper. The authors make a number of choices in order to arrive to the results published. We raise some concerns regarding some of these choices. For example, the authors treat the POWs of the two Andersonville camps as if they were in the same camp. This inflates size of the POWs social network, which is used as the explanatory variable. When we make choices which to us seem more defendable, we cannot, quantitatively and qualitatively, replicate the authors' results.

\section{Summary of the paper}
\textcite{costa2007surviving} try to estimate the effect of quality and size of social networks on ensuring the survival of prisoners of war (POW) in POW camps. To examine the relationship the authors combine two datasets, a longitudinal dataset of Union Army soldiers and a cross-sectional dataset of the POW camp in Andersonville.

The empirical specification for the longitudinal data that measures the effect of the size of the network is as follows:

\begin{equation} \label{eq:ht}
h(t) = \exp ( \beta_1 F + \beta_2 I + \beta_3 M )
\end{equation}

where $h$ is hazard at time $t$, $F$ is the number of friends (=social network), $I$ are individual characteristics and $M$ macro camp conditions. The number of friends is measured as the number of POWs of the same company and varies by month due to deaths or camp transfers. As it might capture the lagged mortality of the group, the authors use as second strategy the initial number of friends as a time-invariant proxy. In addition, they instrument for the number of friends using the number of net camp transfers combined with a dummy indicating whether the POW was transferred to account for omitted variables. 
The specification for the Andersonville data in turn is a probit model:

\begin{align}
Pr(S = 1) &= \Phi (\beta_1 F + \beta_2 F) \label{s1} \\
Pr(S = 1) &= \Phi (\beta_1 F_i + \beta_2 F_{ij} + \beta_3 I | ethnicity = j) \label{s1b} 
\end{align}

where equation \autoref{s1} examines quantity and \autoref{s1b} quality of the social network. The measure for number of friends here is the number of POWs from the same company in Andersonville.

The longitudinal data indicates that POWs had special characteristics. For instance, they were likely to be volunteers, and from companies with more wounded or dead. POWs with more friends were typically US-born and wealthier while Irish and older men had less friends. Concerning survival rates, Kaplan-Meier hazard rates show that POWs captured later (after 1963 when the Union and the Confederate Army stopped exchanging POWs) have lower survival rates. They also show that survival rates were higher for POWs with 10 or more friends. In addition, the rates show that it takes 50 days until there is a divergence in survival between POWs with and without friends. The exponential function estimates that an additional friend leads to a 0.98 times lower mortality hazard. An increase in the number of friends from 0 to 5 (5 to 10), decreases the predicted probability of death from 0.31 to 0.28 (0.26) while a spline specification provides evidence that the effect is greatest for the first two friends. In addition, survival is more likely for POWs that had survived at least one or two months. Factors increasing mortality are the number of men in the camp, low rank, high age, and height among others. Insignificant variables are for instance wealth and marital status. There is no evidence for individual heterogeneity. The use of a Weibull specification indicates no duration dependence (toughen up effects). Interestingly though, a Cox model gives no significant effect of friends which the authors interpret as a result of their loss of power due to the small sample size. Robustness checks, including the IV approach and the effect of friends transferred in, confirms the significant effect of friends. Finally, the cross-sectional results provide evidence that having the same last name, being from the same town for smaller towns, ethnic similarity and having a sergeant or a higher rank in one's network are all important factors in raising survival probabilities which in turn shows that the quality of social networks matters.

In summary, Costa and Kahn are able to show that quantity and quality of social networks mattered for survival in civil war POW camps. Potential explanations for this are the sharing of scarce resources, moral support and protection from other prisoners.

\section{Data}
The raw data for the first part of the paper is extracted from the Union Army dataset readily accessible at \url{http://www.cpe.uchicago.edu}. This dataset draws information from censuses, medical records and military pension records linking military, socioeconomic and health information for nearly $40,000$ white males that served in the Union Army during the Civil War. For the second part, the Andersonville database is available at \url{http://www.itd.nps.gov/cwss} but only as a ``searchable version'', which allows viewing the information for one soldier at a time, and more convenient formats are not readily available.

The datasets provided for the estimations alongside with the publication are already partially processed parts of the raw data available online. No clear guidance is given as to what steps were taken to go from the raw data to the files provided, and matching both things proved to be less than straightforward. In view of this, we decided to proceed with the replication using the files provided by the authors, pointing out along the way the cases where potential problems in them seemed likely given the raw data we observed. Taking those files as starting points, it is possible to reproduce the tables and graphs published in the paper, so in this sense we can say that the data set is complete. Originally, some of the data required by the published .do files was not included in the published data files, but the authors provided it promptly upon our request and added it subsequently to the online files of the publishing journal.

The data and .do files provided do not describe or label many of the variables used, nor does the limited variable description in the Appendix cover all of them. A very thorough description of the variables in the raw data set can be found at its online source, but the match between this and the variables in the files provided is not straightforward or complete, since some of them are already created by the authors from the raw data. The description of the data itself, mainly of the variables that are used most often, is succinct but adequate, with Table 1 and Table 2 presenting some relevant descriptive traits. Despite this, some attention should be drawn to the fact that some of the variables have a relevant share of missing values (only the highest number of observations is reported), a point that will be commented in detail later on.

\section{Stata code and interpretation}
All codes run to their full extent, and produce the estimations presented in the paper, with minor differences that we describe in detail below. The codes themselves are not commented at all and include plenty of estimations and specifications not included in the published paper, as well as unexplained data processing steps. No exporting commands were set up, or log files kept, which might be the root of some of the rounding and significance discrepancies we have found.

In what follows, all reproduced tables are presented with a comment on the differences found (in bold), whenever there were any.

\subsection{Tables}
Table 1: This table presents the means of most variables later used for estimation, dividing observations in POWs and NON POWs, to illustrate ex ante differences in observables between them.

\begin{table}[bt]
\centering
\caption{Characteristics of Soldiers by POW Status}
\label{characteristics}
\begin{tabular}{lllllll}
\hline
                           & All   & Obs.  & Never POW & Obs.  & POW   & Obs. \\
US born                    & 0.745 & 35007 & 0.745     & 31967 & 0.743 & 3040 \\
US born corrected          & 0.747 & 34941 & 0.745     & 31966 & 0.759 & 2975 \\
Irish                      & 0.087 & 34941 & 0.086     & 31966 & 0.102 & 2975 \\
German                     & 0.074 & 34941 & 0.075     & 31966 & 0.062 & 2975 \\
British                    & 0.039 & 34941 & 0.039     & 31966 & 0.033 & 2975 \\
Other                      & 0.054 & 34941 & 0.055     & 31966 & 0.043 & 2975 \\
Artisan                    & 0.2   & 34941 & 0.199     & 31966 & 0.211 & 2975 \\
Farmer corrected           & 0.505 & 34941 & 0.505     & 31966 & 0.508 & 2975 \\
Professional or proprietor & 0.075 & 34941 & 0.077     & 31966 & 0.061 & 2975 \\
Laborer                    & 0.212 & 34941 & 0.212     & 31966 & 0.212 & 2975 \\
Unknown                    & 0.007 & 34941 & 0.007     & 31966 & 0.008 & 2975 \\
Enlisted in 1861           & 0.21  & 35007 & 0.205     & 31967 & 0.263 & 3040 \\
Enlisted in 1861 corrected & 0.21  & 34878 & 0.205     & 31906 & 0.269 & 2972 \\
Enlisted in 1862                  & 0.344   & 34878 & 0.331   & 31906 & 0.487  & 2972 \\
Enlisted in 1863                  & 0.068   & 34878 & 0.066   & 31906 & 0.087  & 2972 \\
Enlisted in 1864                  & 0.256   & 34878 & 0.266   & 31906 & 0.147  & 2972 \\
Enlisted in 1865                  & 0.122   & 34878 & 0.132   & 31906 & 0.009  & 2972 \\
Volunteer                         & 0.909   & 34941 & 0.905   & 31966 & 0.951  & 2975 \\
Height in inches                  & 67.599  & 34941 & 67.589  & 31966 & 67.709 & 2975 \\
Household property income in 1860 & 534.155 & 13769 & 531.734 & 12509 & 558.2  & 1260 \\
Company birthplace fragmentation  & 0.642   & 34941 & 0.644   & 31966 & 0.62   & 2975 \\
Company occupation fragmentation  & 0.559   & 34941 & 0.558   & 31966 & 0.57   & 2975 \\
Fraction company died as non-POWs & 0.135   & 34941 & 0.131   & 31966 & 0.182  & 2975 \\
\end{tabular}
\end{table}

We have added to \autoref{characteristics} the number of observations per variable, not originally included in the paper, in order to highlight the fact that some of them have significant amounts of missing values, something that should be kept in mind later at the estimation stage. In this table, the case in point is Household property income in 1860, which is used in most of the specifications proposed, but as will be discussed for later tables, this is a pattern common to all the variables imported from census data. 
Additionally, considering the number of observations for each of the variables leads to the suggestion of a small correction for three of the dummy variables describing categories: US born, farmer and enlisted in 1861. The initial data file is missing these variables, which are likely to have been dropped as the basis category for estimation purposes in some other use of the data. They were then rebuilt in the .do file from the remaining categories in such a way that generated zeroes for some observations that should have been missing (except for farmer, which is not included at all in the file). For example: as provided by the code, US born has $35,007$ observations, whereas other nationality dummies, all built from the same original variable, have $34,941$. The correction proposed then, is to set these dummies to missing whenever the other categories of their class are missing, which slightly changes their descriptive statistics (for example the mean of US born goes from $0.745$ to $0.747$ when this correction is implemented).

Remaining discrepancies are minor and could be due to some kind of rounding difference. An alternative explanation for these differences is that the table was produced from a slightly different dataset. Considering that even the most populous variables present over 500 observations less than the amount declared in the paper, and the fact that one of the variables (farmer) is not included in the data file at all, we consider the second option to be more likely, which would also explain the minor differences in the means of Artisan, British, Height in inches, etc.

A last and minor comment is that the inclusion of the variable volunteer is surprising, since by sample design the Union Army dataset was restricted to white volunteer infantry regiments. The inclusion of this variable when comparing POWs and non POWs probably responds to the point raised in the paper about POWs not being representative of all soldiers, for example in ideology (p.1471). A variable that reflects that both POWs and non POWs volunteered in similar proportions would suggest homogeneity between POWs and non POWs in this dimension and, hence, increase  external validity. The similarities observed in Table 1 between POWs and non POWs in their mean of volunteer do not reflect that, since they are both close to one by construction. Any individuals not listed as volunteers, either correspond to the very few drafted men, or are actually soldiers that re-enlisted as commissioned officers, but were originally volunteers, as explained in the codebooks for the source data.

\ref{friends}: This table describes the variables later used in the main regressions, focusing in this case only on POWs.  It also presents separately their mean and standard errors for the subsamples of POWs with few (less than 3) and many friends (three or more).

\begin{sidewaystable}
\caption{Characteristics of POWs by Number of Friends in First Two Weeks of Captivity}
\label{friends}
\begin{tabular}{llllllllll}

                                  & \multicolumn{3}{l}{All} & \multicolumn{3}{l}{Few friends ($<3$)} & \multicolumn{3}{l}{Many friends ($>=3$)} \\
                                  & Mean                    & Std. err.                            & No. obs.                               & Mean & Std. err. & No. obs. & Mean & Std. err. & No. obs. \\
US born                           & 0.731                   & 0.444                                & 1976                                   & 0.708 & 0.455 & 990 & 0.754 & 0.431 & 986 \\
Irish                             & 0.12                    & 0.325                                & 1976                                   & 0.115 & 0.319 & 990 & 0.125 & 0.331 & 986 \\
German                            & 0.063                   & 0.243                                & 1976                                   & 0.081 & 0.273 & 990 & 0.045 & 0.207 & 986 \\
British                           & 0.034                   & 0.182                                & 1976                                   & 0.042 & 0.202 & 990 & 0.026 & 0.16 & 986 \\
Other                             & 0.05                    & 0.217                                & 1976                                   & 0.053 & 0.223 & 990 & 0.047 & 0.211 & 986 \\
Artisan                           & 0.209                   & 0.407                                & 1976                                   & 0.214 & 0.41 & 990 & 0.204 & 0.403 & 986 \\
Farmer                            & 0.421                   & 0.494                                & 1976                                   & 0.44 & 0.497 & 990 & 0.401 & 0.49 & 986 \\
Professional or propietor         & 0.07                    & 0.256                                & 1976                                   & 0.077 & 0.266 & 990 & 0.064 & 0.245 & 986 \\
Laborer                           & 0.266                   & 0.442                                & 1976                                   & 0.243 & 0.429 & 990 & 0.288 & 0.453 & 986 \\
Unknown                           & 0.034                   & 0.182                                & 1976                                   & 0.025 & 0.157 & 990 & 0.044 & 0.204 & 986 \\
Enlisted in big city              & 0.475                   & 0.5                                  & 1976                                   & 0.493 & 0.5 & 990 & 0.457 & 0.498 & 986 \\
Household property income in 1860 & 576.475                 & 1855.699                             & 817                                    & 483.609 & 1084.987 & 391 & 661.711 & 2348.676 & 426 \\
Age when captured                 & 25.96                   & 7.056                                & 1966                                   & 26.645 & 7.415 & 990 & 25.265 & 6.604 & 976 \\
Height when captured                 & 171.576 & 6.612 & 1925 & 171.79 & 6.384 & 974 & 171.356 & 6.833 & 951 \\
Commisoned or noncommisoned officer  & 0.076   & 0.266 & 1976 & 0.083  & 0.276 & 990 & 0.07    & 0.255 & 986 \\
Married in 1860                      & 0.257   & 0.437 & 817  & 0.276  & 0.448 & 391 & 0.239   & 0.427 & 426 \\
Wounded 10 days before capture       & 0.118   & 0.323 & 1976 & 0.148  & 0.356 & 990 & 0.088   & 0.284 & 986 \\
Fraction company dead before capture & 0.089   & 0.064 & 1976 & 0.093  & 0.07  & 990 & 0.085   & 0.057 & 986 \\
Company birthplace fragmentation     & 0.542   & 0.2   & 1923 & 0.547  & 0.193 & 985 & 0.536   & 0.206 & 938 \\
Company occupational fragmentation   & 0.596   & 0.173 & 1923 & 0.586  & 0.173 & 985 & 0.607   & 0.172 & 938 \\
\end{tabular}
\end{sidewaystable}

This table is almost exactly reproduced, except for the minor differences set in bold, which appear to be rounding differences (and a typo in the case of Laborer). The number of observations stated on the table in the paper (1923) corresponds to the last variable in this case as well, even though, as mentioned for the previous table, it is relevant to keep in mind that some of the other variables of interest show a considerably lower count of observations, as can be seen in the added columns. For this table, this is the case of variables Married in 1860 and Household property income in 1860, which are both variables from the 1860 census.
The main concern regarding this table is the process of organizing the data, particularly one step in which a merge to the file ``transfer.dta'' is conducted. This file contains the prisoner identifier variable, month, year, and a dummy variable reflecting whether the prisoner was transferred (transferin). This variable is not needed for \autoref{friends}, and we can only speculate that it was incorporated, as some other steps were, in an attempt to exactly reproduce the dataset used in \autoref{corrected}. The problem in this case is that the file ``transfer.dta'' contains duplicate observations, which in turn duplicates observations in the main data set via the merge, artificially inflating the number of observations in the data used for \autoref{friends}\footnote{Duplicate observations per month-year-prisoner are natural in the context of \autoref{corrected}, because in there capture time is measured per fortnight. Including additional observations because of this in a stage of data description such as this one is hardly appropriate.}.

If these duplicate observations are eliminated from the ``transfer.dta'' file before the merge, the statistics presented do not seem to defer significantly, and certainly neither do the conclusions drawn from them, but the number of observations in the subsample considered is significantly lower. This exercise is presented in the table below.

>>>>

\begin{refsection}
\chapter{Bitcoin and Remitances: The Potential of `Stupid Phones'}
\label{btc}
In the 2009 whitepaper \parencite{nakamoto2008bitcoin} called \citetitle{nakamoto2008bitcoin} an anonymous individual or collective referred to a Satoshi Nakamoto (the Japanese equivalent of John Smith) describes an online currency called Bitcoin (BTC). 
This currency -the first of a group called cryptocurrencies- has the following economics features:
